---
title: "migrate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{migrate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(width = 999)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(migrate)
```

## Using `migrate`

This package is intended to serve as a set of tools to help convert credit risk data at two time points into traditional credit state migration (aka, "transition") matrices. At a higher level, `migrate` is intended to help an analyst understand how risk moved in their credit portfolio over a time interval.

## Background

One of the more difficult aspects of making a credit state migration matrix in R (or Python, for that matter) is the fact that the output doesn't satisfy the structure of a traditional data frame object. Rather, the output needs to be a *matrix*, which is a data structure that R does support. In the past, there has been difficulty converting a matrix to something more visual-friendly. More recently, however, tools like the [kableExtra](https://cran.r-project.org/package=kableExtra) and [gt](https://cran.r-project.org/package=gt) packages allow us to present visually appealing output that extends the structure of a data frame. Using the matrix-style output of `migrate`'s functions with a visual formatting package such as the two mentioned above will hopefully help analysts streamline the presentation of their credit portfolio's state migration matrices to an audience.

## Getting Started

If you haven't done so already, first install `migrate` with the instructions in the [README section](https://github.com/mthomas-ketchbrook/migrate#Installation).

First, load the package & the mock dataset (as a data frame) using `library()`

```{r load, eval = FALSE}
library(migrate)
```

The package has a built-in mock dataset, which can be loaded into the environment like so:

```{r data, results = 'hide'}
data("mock_credit")

head(mock_credit[order(mock_credit$customer_id), ])   # sort by 'customer_id'
```

```{r data_tbl, echo = FALSE}
head(mock_credit[order(mock_credit$customer_id), ]) %>% 
  knitr::kable(row.names = FALSE)
```

Note that an important feature of the `mock_credit` dataset is that there are exactly two (2) unique values in the `date` column variable; if the `time` argument passed to `migrate` has more than two (2) unique values, the function will throw an error.

```{r dates}
unique(mock_credit$date)
```

To summarize the migration within the data, use the `migrate()` function

```{r migrate}
migrated_df <- migrate(
  data = mock_credit, 
  id = customer_id, 
  time = date, 
  state = risk_rating, 
)

head(migrated_df)
```

To create the state migration matrix, use the `build_matrix()` function
```{r matrix}
build_matrix(migrated_df)
```

Or, to do it all in one shot, use the `%>%`
```{r pipe} 
mock_credit %>% 
  migrate(
    id = customer_id, 
    time = date, 
    state = risk_rating, 
    metric = principal_balance, 
    percent = FALSE, 
    verbose = FALSE
  ) %>% 
  build_matrix(
    state_start = risk_rating_start, 
    state_end = risk_rating_end, 
    metric = principal_balance
  )
```
