---
title: "migrate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{migrate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(width = 999)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Using `migrate`

This package is intended to serve as a set of tools to help convert credit risk data at two time points into traditional credit state migration (aka, "transition") matrices. At a higher level, `migrate` is intended to help an analyst understand how risk moved in their credit portfolio over a time interval.

## Background

One of the more difficult aspects of making a credit state migration matrix in R (or Python, for that matter) is the fact that the output doesn't satisfy the structure of a traditional data frame object. Rather, the output needs to be a *matrix*, which is a data structure that R does support. In the past, there has been difficulty converting a matrix to something more visual-friendly. More recently, however, tools like the [kableExtra](https://cran.r-project.org/package=kableExtra) and [gt](https://cran.r-project.org/package=gt) packages allow us to present visually appealing output that extends the structure of a data frame. Using the matrix-style output of `migrate`'s functions with a visual formatting package such as the two mentioned above will hopefully help analysts streamline the presentation of their credit portfolio's state migration matrices to an audience.

## Getting Started

If you haven't done so already, first install `migrate` with the instructions in the [README section](https://github.com/mthomas-ketchbrook/migrate#Installation).

First, load the package & the mock dataset (as a data frame) using `library()`

```{r load}
library(migrate)
data("mock_credit")
```

We can get a look at the data using `dplyr::glimpse()`

```{r glimpse, warning = FALSE, message = FALSE}
library(dplyr)
glimpse(mock_credit)
```

Note that an important feature of the dataset is that there are exactly two (2) unique values in the `date` column variable

```{r dates}
unique(mock_credit$date)
```

To summarize the migration within the data, use the `migrate()` function

```{r migrate}
migrated_df <- migrate(
  data = mock_credit, 
  date = date, 
  rating = risk_rating, 
  metric = principal_balance
)

head(migrated_df)
```

To create the state migration matrix, use the `build_matrix()` function
```{r matrix} 
build_matrix(migrated_df)
```

Or, to do it all in one shot, use the `%>%`
```{r pipe, eval = FALSE} 
mock_credit %>% 
  migrate(
    data = mock_credit, 
    date = date, 
    rating = risk_rating, 
    metric = principal_balance
  ) %>% 
  build_matrix()
```
