---
title: "migrate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{migrate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Using `migrate`

This package is intended to serve as a set of tools to help convert credit risk data at two time points into traditional credit state migration (aka, "transition") matrices. At a higher level, `migrate` is intended to help an analyst understand how risk moved in their credit portfolio over a time interval.

## Background

One of the more difficult aspects of making a credit state migration matrix in R (or Python, for that matter) is the fact that the output doesn't satisfy the structure of a traditional data frame object. Rather, the output needs to be a *matrix*, which is a data structure that R does support. In the past, there has been difficulty converting a matrix to something more visual-friendly. More recently, however, tools like the [kableExtra]() and [gt]() packages allow us to present visually appealing output that extends the structure of a data frame. Using the matrix-style output of `migrate`'s functions with a visual formatting package such as the two mentioned above will hopefully help analysts streamline the presentation of their credit portfolio's state migration matrices to an audience.

## Getting Started

If you haven't done so already, first install `migrate` with the instructions in the [README section](https://github.com/mthomas-ketchbrook/migrate#Installation)
```{r install}
# Install the lastest version on CRAN
install.packages("")


```


```{r setup}
library(migrate)
```
